From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-18 paper.tex
Results in table are actually with-me-ness values
First pass on section 3-Experimental validation, including a new, tikz-based system overview

@@

@@

\section{Visual Attention Assessment}

@@

As presented above, we use the (6D) head pose as an approximation of the
actual gaze direction, and we approximate from here the participant's field of
attention. The assumption that such an approximation of the field of attention
allows to derive the actual focus of attention is the core hypothesis that we
ought to validate.

This section presents such an experimental validation, set in the context of a
real child-robot interaction scenario where the robot tries to engage a child in
handwriting tasks using learning by teaching paradigm.

\paragraph{Real-Time Assessment of Visual Focus of Attention}

The participants' VFoA acquisition is the only dependent variable. This includes
their gaze times and patterns of looking or not looking to the targets. This
data extracted from an RGB camera was also coded post-hoc to quantify some of
these measurements.

We predict that there will be no generalizable pattern of participant's VFoA
behavior for different scenarios such as multiple subjects, necessitating in
situ VFoA estimation capabilities to enable robots to adapt in the specific
scenarios. However, we believe that off-line pattern identification can lead to
a better understanding of the children progression as well as engagement level
assessment in face-to-face long term interactions.\fixme{I did not understand
this paragraph}

\fixme{Underline difference between field of attention and focus of attention}

\fixme{State the hypothesis: head-pose estimation provides a good-enough
approximation of the gaze direction, and in turn of the field of attention, and
in turn of the focus of attention}

@@

The target of this study were 6 children with ages compressed between 5 and 6
years old. The experiments were developed individually and consisted on a
writing activity by turns and one story telling performed by the robot at a
given time of the interaction. All this performed in a over all 18 minutes
interaction average time.

Figure~\ref{fig:realSetup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
            nao} robot.  The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting its letters on the tablet
            with a stylus. An adult (either a therapist or an experimenter,
            depending on the studies), remains next to the child to guide the work. 
            A second tablet allows to choose demonstrations and the camera captures the subject's face.}
    \label{fig:realSetup}
\end{figure}

A tactile tablet (with a custom application) is used for both the robot and the
child to write: during a typical round, the child requests the robot to write
something (a single letter, a number or a full word), and push the tablet
towards the robot, the robot writes on the tablet by gesturing the writing (but
without actually physically touching the tablet), the child then pull back the
tablet, corrects the robot's attempt by writing him/herself on top or next to
the robot's writing, and ``send'' his/her demonstration to the robot by pressing
a small button on the tablet. The robot ``learns'' from this demonstration and
tries again.

Since the child is assumed to take on the role of the teacher, we had to ensure
(s)he would be able to manage by him/herself the turn-taking and the overall
progression of the activity (moving to the next letter or word). In our design,
the turn-taking relies on the robot prompting for feedback once it is done with
its writing (simple sentences like ``What do you think?''), and pressing on a
small robot icon on the tablet once the child has finished correcting. In our
experiments, both were easy to grasp for children.

Implementing such a system raises several challenges that are discussed in
detail in \cite{Hood:2015}.

An RGB camera has been used to acquire images of 640x480 pixels at 25 fps. The
camera was located at 5 cm from the center of Nao's feet and its holder was tied
to a wood base for stability purposes. Thus, the camera's objective was 9 cm
above and $ 40\degree $ inclination with respect to the surface of the table to
maximize the detection during the writing time. However, Nao's camera could have
been used, always considering the \textit{tf} frame matrix update according to
the head orientation.

The subjects were typically located 50 cm away from the robot with the first
tablet in front and the second one 30 cm to the left of the first one. In the
same way, the experimenter was located around 60 cm to the left of the subject.
Finally, two observers were located far away from the interaction field to
manually assess the state of the interaction.

Figure \ref{drawSetup} allows us to intuitively identify several focuses of
attention along the interaction such as the two tablets, the robot, the
experimenter and the observers located in the diagonal.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\columnwidth]{experimental_setup}
    \caption{\small \textbf{Experimental setting}. The areas of interest (AOI)
    of potential focuses of attention are circled in red.}
    \label{drawSetup}
\end{figure}

\subsection{System Implementation}
\label{sec:system}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{system}
    \caption{\small System overview showing the most relevant components involved in the estimation.}
    \label{system}
\end{figure}

, to detect and extract the skeleton
of the subject's face and then, use the obtained features to calculate the
head pose orientation estimation (see figure \ref{system}). In this way, we can
estimate what the subject was looking at and obtain an \textit{rviz} 3D scene
representation as it is executed as a ROS (Robot Operating System) node that can
be seen in figure \ref{rviz}.

@@

\subsubsection{From Face Detection to Focus of Attention} \label{fromFaceTo}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\columnwidth]{bitmap}
    \caption{\small Cosine similarity matrix between real and cap-tured VFoA
    using a 10 sec. window size computation. The correspondent numbers are shown
    in figure \ref{fig:realCaptured}.}
    \label{fig:bitmap}
\end{figure}

@@

\begin{table}[h!]
    \centering
    \caption{\textbf{Levels of with-me-ness}. For each subject, the with-me-ness level is reported over the whole interaction.}
    \begin{tabular}{p{2cm}cccccccc}
        \toprule
        \textbf{Subject} & 1 & 2 & 3 & 4 & 5 & 6 & Avg. & SD \\
        \midrule
        %Ground-truth \vs expected (\%) & 92.6 & 88.8  & 95.8 & 93.1 & 96.5 & 89.6 & 92.7 & \\ \hline
        $\mathcal{W}_{groundtruth}$ &  & & & & & & & \\ 
        {\it Cronbach's $\alpha$} &  & & & & & & & \\
        \midrule
        $\mathcal{W}_{robot}$ &  & & & & & & & \\
        {\it Cronbach's $\alpha$} &  & & & & & & & \\
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table}

>>>>

\usepackage{hyperref}

@@

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning, calc}

@@

\section{Visual Attention Assessment}
\label{sec:vfoa}

@@

As presented above, we use the (6D) head pose as an approximation of the
actual gaze direction, and we approximate from here the participant's field of
attention. The assumption that such an approximation of the field of attention
allows to derive the actual focus of attention forms the hypothesis that we
want to validate experimentally. The experiment involves child-robot
interactions in the context of handwriting remediation.
This section details the experimental procedure and presents our results.

@@

The experiment, part of the CoWriter project~\cite{Hood:2015}, involves a robot
which tries to engage a child in handwriting tasks using a \emph{learning by teaching}
paradigm (\ie the child is the teacher, and he/she attempts to improve the
robot's handwriting). A tactile tablet is used as writing support.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\columnwidth]{experimental_setup}
    \caption{\small \textbf{Experimental setup}: face-to-face interaction with a {\sc
            nao} robot. The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting its letters on the tablet
            with a stylus. The facilitator remains next to the child to guide the work. 
            The secondary tablet allows the child to tell the robot what to
            write. The areas of interest -- corresponding to potential target of
            attention -- are circled in red with their name.}
    \label{fig:setup}
\end{figure}

Figures~\ref{fig:setup} and~\ref{fig:realSetup} illustrate the experimental setup: a face-to-face
child-robot interaction with an (autonomous) Aldebran's {\sc nao} robot, in
presence of a facilitator (one of the researchers).

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Picture of the interaction with one of the children.}
    \label{fig:realSetup}
\end{figure}

The dependent variable is the measurement of the participants' VFoA, assessed in
terms of what are the attentional targets of the child over time. The face of the child
is acquired through a fixed webcam (Logitech {\sf c920}), placed on the table
(see Figure~\ref{fig:setup}), and the attentional targets are then computed as
presented in section~\ref{sec:vfoa}.

%An RGB camera has been used to acquire images of 640x480 pixels at 25 fps. The
%camera was located at 5 cm from the center of Nao's feet and its holder was tied
%to a wood base for stability purposes. Thus, the camera's objective was 9 cm
%above and $ 40\degree $ inclination with respect to the surface of the table to
%maximize the detection during the writing time. However, Nao's camera could have
%been used, always considering the \textit{tf} frame matrix update according to
%the head orientation.

The subjects were typically located 50 cm away from the robot with the primary
(writing) tablet in front and the secondary one 30 cm to the left of the first
one.  The facilitator was located about 60 cm to the left of the subject.
Finally, two observers were located further away from the interaction field.
Figure~\ref{fig:setup} indicates accordingly the location of main areas of
interest (the two tablets, the robot, the facilitator and the observers).

Six children (ages 5 to 6\fixme{boys? girls?}) were enrolled for this study.
The study took place at school, in a separate room (the computer
lab). The participants were chosen by the teacher, and would come one after the
other for a face-to-face interaction with a {\sc Nao} robot (average duration:
18 min, $SD:$ \fixme{stddev?}).

The experiment is organized in rounds of writing: during a typical round, the
child requests the robot to write something (a single letter, a number or a full
word), and pushes a tactile tablet (equipped with a custom application that we
developed) towards the robot; the robot writes on the tablet by mimicking the
writing gesture (but without actually physically touching the tablet); the child
then pull back the tablet, corrects the robot's attempt by writing him/herself
on top or next to the robot's writing, and ``send'' his/her demonstration to the
robot by pressing a small button on the tablet. The robot learns from this
demonstration and tries again. The child continues the turn-taking until he/she
decides to train the robot on another word. Details on the rationale and
implementation of this study can be found in~\cite{Hood:2015}.

\fixme{how many rounds for each child?}

Besides, once per interaction, the robot did interrupt the handwriting task to
propose to tell a funny story. The story-telling lasted in average \fixme{how
long?} min, and the turn-based hand-writing task continued afterwards. The
intended purpose of the story-telling episode was to locally trigger a different
set of gaze behaviors.

\subsection{System Implementation}

\label{sec:system}

\begin{figure}[h!]
    \centering

\resizebox{0.9\linewidth}{!}{%

\begin{tikzpicture}[
    >=latex,
    node distance=2cm,
    every edge/.style={draw, very thick},
    redarrow/.style={fill=Red, draw=Red},
    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]

%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};

    \node at (0,0) [cmpt, fill=Green!50] (tf) {tf};
    \node [cmpt, fill=YellowOrange!50, left=of tf] (attention) {attention tracking};
    \node [above=0.6cm of attention, text width=2.7cm,align=center] (input) {\tt /image \\ /camera\_info \\ dlib {\sf faces model} \\ {\sf 3D face template}};
    \node [cmpt, right=of tf] (static) {static pose publishers};
    \node [above=0.6cm of static, text width=2.7cm,align=center] (staticposes) {\sf locations of \\areas of interest};
    \node [cmpt, above=0.6cm of tf] (pose) {robot state publisher};
    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tf] (focus) {focus estimation};
    \node [below=0.6cm of focus] (output) {\tt /frames\_in\_focus};

    \path (staticposes) edge [->] (static);
    \path (static) edge [->] (tf);
    \path (input) edge [->] (attention);
    \path (attention) edge [->] (tf);
    \path (pose) edge [->]  (tf);
    \path (tf) edge [->]  (focus);
    \path (focus) edge [->]  (output);

\end{tikzpicture}
}

    \caption{\small ROS nodes involved in the VFoA estimation (yellow nodes were developed for this work).}
    \label{fig:system}
\end{figure}

The experiment was carried with an Aldebaran's {\sc Nao} robot, using ROS as a
middleware to build the attention estimation pipeline
(Figure~\ref{fig:system}). Head pose estimation builds on the {\tt dlib} and
OpenCV libraries, while pose transformations are handled by the {\sc tf}
library. The implementation is open-source and available at
\url{https://github.com/chili-epfl/attention-tracker}.

Please refer to~\cite{Hood:2015} for the implementation of the hand-writing
activity.

@@

\subsubsection{From Face Detection to Focus of Attention} \label{fromFaceTo}
(note that the focus of attention
was separately hand-coded, {\it post-hoc}, from the video recording of the
interaction).

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\columnwidth]{bitmap}
    \caption{\small Cosine similarity matrix between ground-truth and robot-measured VFoA
    using a 10 sec. window size. The correspondent numbers are shown
    in figure \ref{fig:realCaptured}.}
    \label{fig:bitmap}
\end{figure}

@@

\begin{table}[h!]
    \centering
    \caption{\textbf{Levels of with-me-ness}. For each subject, the with-me-ness level is reported over the whole interaction, either based on the ground-truth (\ie \emph{ground-truth with-me-ness}), or based on the focus of attention measured by the robot.}
    \begin{tabular}{p{1cm}cccccccc}
        \toprule
        Subject & 1 & 2 & 3 & 4 & 5 & 6 & Avg. & SD \\
        \midrule
        $\mathcal{W}_{g.truth}$ & 92.6 & 88.8  & 95.8 & 93.1 & 96.5 & 89.6 & 92.7 & 3.1 \\ 
        %{\it Cronbach's $\alpha$} &  & & & & & & & \\
        %\midrule
        $\mathcal{W}_{robot}$ &  & & & & & & & \\
        %{\it Cronbach's $\alpha$} &  & & & & & & & \\
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table}
